Towards a Machine Learning approach for Aggregate Computing

### 2021-11-28 paper-2021-woa-ac-ml.tex
On motivation

\subsection{Motivation}
Currently, Aggregate computing has a problem in the ability to generalise collective behaviours that are capable of working well in different scenarios. 
 For example, a well-tested aggregate application in a static network could lead to a non-stable behaviour in a dynamic network. 
 Typically, to alleviate this problem, designers create ad-hoc building blocks that cope with certain environmental aspects (such as the ULT block~\cite{DBLP:conf/saso/AudritoCDV17}). 
 But they are still context-specific solutions.

A novel way to deal with this problem is to leave the low-level aspects of AC to Machine Learning algorithms, 
 training the collective in different environments and thus building general behaviours that are capable of adapting to many circumstances.

In this case, when Reinforcement Learning is applied to several agents, we speak of Multi-Agent Reinforcement Learning (MARL)~\cite{DBLP:journals/aamas/Hernandez-LealK19}.
 AC has a typical cooperative and homogeneous learning setting: 
 each agent collaborates to achieve a common goal guided by the same reward function.
%
In the cooperative and homogeneous systems, several works have tried to apply the algorithms used in the single-agent context to the multi-agent one --- 
 the so-called Independent MARL~\cite{DBLP:journals/tsmc/BusoniuBS08}. 
%
However, several issues emerge, such as non-stationarity, multi-agent credit assignment, scalability and partial observability of the system. 
 Therefore, several heuristics have tried to solve these problems.
%
Among many Hysteric Q-Learning~\cite{hysteretic-q} uses an optimistic approach in which more weight is given to positive actions than negative ones.

Another way to handle multi-agent learning is to use so-called \textit{Centralised Training and Decentralised Execution} (CTDE) algorithms. 
 In this case, a central agent collects the experiences of the various agents and then distils a policy that each agent can execute locally in a typical actor-critic setting. 
%
During the learning phase, the problems of partial observability and non-stationarity are reduced, since the centralised learner has a global view. 
 However, the problem of scalability is typically not addressed.

>>>>

\subsection{Motivation}

Consider the problem of associating
 to each node of the network
 a value
 representing the estimation of the minimum distance
 from that node to the closest \emph{source} node (see \Cref{fig:hop-count}).
%
This is indeed a computational field 
 and it is known as a \emph{gradient}~\cite{DBLP:conf/saso/AudritoCDV17}
 and there exist aggregate algorithms
 for computing it in such a way that
 the field automatically adapts 
 to changes induced by mobility
  and addition/removal of sources.
%
The gradient is a key building block 
 for building more complex self-organising behaviour~\cite{DBLP:journals/jlap/ViroliBDACP19}.

%Currently, aggregate computing has a problem in the ability to generalise collective behaviours that are capable of working well in different scenarios. 
%%
% For example, a well-tested aggregate application in an environment with limited dynamics could lead to a non-stable behaviour under continuous perturbations.
%% 
% Typically, to alleviate this problem, designers create ad-hoc building blocks that cope with certain environmental aspects (such as the ULT block~\cite{DBLP:conf/saso/AudritoCDV17}). 
% But they are still context-specific solutions.
However, in practice, 
 specific gradient algorithms
 may be developed
 to achieve different trade-offs 
 in terms of reactivity to change and stability~\cite{DBLP:conf/saso/AudritoCDV17},
 which make them more or less suitable depending on 
 the environmental and network dynamics at hand.
%
The problem is that manually crafting such algorithms
 is not an easy task, for various heuristics and assumptions typically have to be considered.
%
To address this problem,
 we propose to use machine learning techniques
 in order to learn ``good'' local decision-making.
% 
That is, the idea is to leave the low-level aspects of AC to Machine Learning algorithms, 
 training the collective in different environments and thus building general behaviours that are capable of adapting to many circumstances.

In this case, when Reinforcement Learning is applied to several agents, we speak of Multi-Agent Reinforcement Learning (MARL)~\cite{DBLP:journals/aamas/Hernandez-LealK19}.
 AC has a typical cooperative and homogeneous learning setting: 
 each agent collaborates to achieve a common goal guided by the same reward function.
%
In the cooperative and homogeneous systems, several works have tried to apply the algorithms used in the single-agent context to the multi-agent one --- 
 the so-called Independent MARL~\cite{DBLP:journals/tsmc/BusoniuBS08}. 
%
However, several issues emerge, such as non-stationarity, multi-agent credit assignment, scalability and partial observability of the system. 
 Therefore, several heuristics have tried to solve these problems.
%
Among many, Hysteric Q-Learning~\cite{hysteretic-q} uses an optimistic approach in which more weight is given to positive actions than negative ones.
%
Another way to handle multi-agent learning is to use so-called \textit{Centralised Training and Decentralised Execution} (CTDE) algorithms. 
 In this case, a central agent collects the experiences of the various agents and then distils a policy that each agent can execute locally in a typical actor-critic setting. 
%
During the learning phase, the problems of partial observability and non-stationarity are reduced, since the centralised learner has a global view. 
 However, the problem of scalability is typically not addressed.
